---
title: "Differential expression explorer with functional enrichment"
author: "Marek Gierlinski"
package: volcenrich
output:
  BiocStyle::html_document:
    toc_float: true
abstract: |
  This R package, designed for interactive exploration and analysis
    of differential expression or abundance data, integrates Shiny-based
    visualization tools with rapid functional enrichment analysis. Users
    can dynamically select genes from volcano or MA plots and instantly access
    enriched GO, KEGG, and Reactome pathway analyses, facilitating a deeper
    understanding of gene functions and biological pathways. The package
    supports customized plotting, feature information retrieval, and handles
    large datasets efficiently, making it an ideal tool for genomic
    researchers and bioinformaticians seeking to uncover the biological
    significance behind differential expression patterns.
vignette: >
  %\VignetteIndexEntry{Fast functional enrichment}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
options(width = 72)
```

# Purpose

# Installation

*fenr* can be installed from *Bioconductor* by using:

```{r bioconductor_install, eval=FALSE}
if (!require("remotes", quietly = TRUE))
  install.packages("remotes")

remotes::install_github("bartongroup/volcenrich")
```

# Overview

The input data for *volcenrich* consists of 5 objects: `de`, `data`, `metadata`, `features` and `fterms`.

## Differential expression results

Argument name `de`. It is a data frame (or tibble) with the following columns:

  - `id` (character) - feature identifier,
  - `log_fc` (numeric) - log-fold change,
  - `expr` (numeric) - mean expression or abundance,
  - `p_value` (numeric) - p-value from statistical tests,
  - `fdr` (numeric) - false discovery rate, p-value corrected for multiple tests,
  - `contrast` (factor) - name of the contrast; this data frame can contain results from multiple contrasts.







# Session info {.unnumbered}

```{r session_info, echo=FALSE}
sessionInfo()
```
